---
title: "Cylistics"
author: "Sourabh Kumar"
date: "11/16/2021"
output: html_document
---


##Load the library
```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(data.table)
library(ggthemes)
library(readr)
```
## Get the directories and set as well
```{r}
getwd()
setwd( "C:/Users/subha/Desktop/Case Study/casestudy")
```
##Load the files 
```{r}
apr20<- read_csv("C:/Users/subha/Desktop/Case Study/google/202004-divvy-tripdata.csv")
may20 <- read_csv("C:/Users/subha/Desktop/Case Study/google/202005-divvy-tripdata.csv")
june20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202006-divvy-tripdata.csv")
july20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202007-divvy-tripdata.csv")
aug20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202008-divvy-tripdata.csv")
sep20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202009-divvy-tripdata.csv")
oct20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202010-divvy-tripdata.csv")
nov20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202011-divvy-tripdata.csv")
dec20 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202012-divvy-tripdata.csv")
jan21 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202101-divvy-tripdata.csv")
feb21 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202102-divvy-tripdata.csv")
mar21 <-  read_csv("C:/Users/subha/Desktop/Case Study/google/202103-divvy-tripdata.csv")

```
##Check the column names of files
```{r echo=TRUE}
colnames(apr20)
colnames(may20)
colnames(june20)
colnames(july20)
colnames(aug20)
colnames(sep20)
colnames(oct20)
colnames(nov20)
colnames(dec20)
colnames(jan21)
colnames(feb21)
colnames(mar21)
```
### Takes list, vector or data frame as input and gives output in vector or matrix with the help of sapply().
```{r echo=TRUE}
sapply(apr20, typeof)
sapply(may20,typeof)
sapply(june20, typeof)
sapply(july20, typeof)
sapply(aug20, typeof)
sapply(sep20, typeof)
sapply(oct20, typeof)
sapply(nov20, typeof)
sapply(dec20, typeof)
sapply(jan21, typeof)
sapply(feb21, typeof)
sapply(mar21, typeof)
```
##Change the data types
```{r}
dec20 <- mutate(dec20, start_station_id = as.double(start_station_id),
end_station_id = as.double(end_station_id))
jan21 <- mutate(jan21, start_station_id = as.double(start_station_id),
end_station_id = as.double(end_station_id))
feb21 <- mutate(feb21, start_station_id = as.double(start_station_id),
end_station_id = as.double(end_station_id))
mar21 <- mutate(mar21, start_station_id = as.double(start_station_id),
end_station_id = as.double(end_station_id))
```
##Re-check the data types
```{r echo=TRUE}
is.double(dec20$start_station_id)
is.double(dec20$end_station_id)
is.double(jan21$start_station_id)
is.double(jan21$end_station_id)
is.double(feb21$start_station_id)
is.double(feb21$end_station_id)
is.double(mar21$start_station_id)
is.double(mar21$end_station_id)
```
##Bind all the files into one.
```{r}
all_bike_trips <- bind_rows(apr20,may20,june20,july20,aug20,sep20,oct20,nov20,
dec20,jan21,feb21,mar21)
```
##Check the new file
```{r echo=TRUE}
summary(all_bike_trips)
dim(all_bike_trips)
```
## Rename some column names for better readability
```{r}
all_bike_trips <- all_bike_trips %>% rename(
type_of_ride = rideable_type,
start_time = started_at,
end_time = ended_at,
type_of_member = member_casual)
```
## Re-check the file
```{r echo=TRUE}
str(all_bike_trips)
```
## Add a column called ride length to find total trip duration in minutes
```{r}
all_bike_trips$ride_length <- (as.double(difftime(all_bike_trips$end_time,
all_bike_trips$start_time))) / 60
```
```{r echo=TRUE}
glimpse(all_bike_trips)
```
##Check null values
```{r}
sum(all_bike_trips$ride_length < 0)
```
```{r}
all_bike_trips_v2 <- all_bike_trips[!(all_bike_trips$ride_length < 0),]
all_bike_trips_v2<- all_bike_trips_v2[!((all_bike_trips_v2$start_station_name %like% "TEST" |
all_bike_trips_v2$start_station_name %like% "test" |
all_bike_trips_v2$start_station_name %like% "Test")),]
```
```{r}
all_bike_trips_v2 <- all_bike_trips_v2 %>% select(-c(start_lat:end_lng))
```
##Check the data
```{r echo=TRUE}
glimpse(all_bike_trips_v2)
```
```{r}
all_bike_trips_v2$date <- as.Date(all_bike_trips_v2$start_time) ### The default is yyyy/mm/dd
```
```{r}
all_bike_trips_v2$month <- format(as.Date(all_bike_trips_v2$date), "%b_%Y")
```
```{r}
all_bike_trips_v2$day <- format(as.Date(all_bike_trips_v2$date), "%d") ###dd
```
```{r}
all_bike_trips_v2$year <- format(as.Date(all_bike_trips_v2$date), "%Y") ###yyyy
```
```{r}
all_bike_trips_v2$day_of_week <- format(as.Date(all_bike_trips_v2$date), "%a") ###Sun, Mon
```
## We have to use POSIXct object when dealing with times
## We are first striping the date and just leaving the time, this get formatted as a chr
```{r}
all_bike_trips_v2$time <- format(as.POSIXct(all_bike_trips_v2$start_time), format = "%H:%M")
```
## We then the time column into format - POSIXct, however we still will get a date, but this will just be our current date, so if we want to use this time column we just need to reformat it when used in a visualizations
```{r}
all_bike_trips_v2$time <- as.POSIXct(all_bike_trips_v2$time, format = "%H:%M")
```
## Check our data set for NA (null) values
```{r echo=TRUE}
sum(is.na(all_bike_trips_v2) == FALSE) ## Returns sum of non NA values
sum(is.na(all_bike_trips_v2) == TRUE) ## Returns sum of NA values
```
## 1057208 (total NA values) is 2.03 % of 52132260 (total values)
```{r echo=TRUE}
summary(all_bike_trips_v2)
```

#From here We can visualize the data with the help of tableau
